Integrating Common Ground and Informativeness in Pragmatic Word Learning

### 2019-01-09 papers/cogsci2019/cogsci2019.tex
first draft CogSci paper

---
title: "Integrating Common Ground and Informativeness in Pragmatic Word Learning"
bibliography: library.bib
csl: apa6.csl
document-params: "10pt, letterpaper"

author-information: > 
   \author{{\large \bf Manuel Bohn} \\ \texttt{bohn@stanford.edu} \\ Psychology, Stanford University \\ LFE, Leipzig University 
    \And {\large \bf Michael Henry Tessler} \\ \texttt{tessler@mit.edu} \\ Department of Brain and Cognitive Sciences \\ Massachusetts Institute of Technology
    \And {\large \bf Michael C. Frank} \\ \texttt{mcfrank@stanford.edu} \\ Department of Psychology \\ Stanford University}

abstract: 
    "Pragmatic inferences are an integral part of language comprehension and learning. To recover the intended meaning of an utterance, listeners need to balance and integrate different sources in information. In a series of experiments, we studied how listeners integrate general expectations about speakers with expectations that are specific to the interactional history with  a particular speaker. We used a Bayesian cognitive model to formalize the integration process. In Experiment 1 and 2 we replicated findings showing that listeners make inferences based on general and speaker specific expectations. Using the empirical measurements from these experiments, we generated model predictions about how they should be integrated. We tested these predictions in Experiment 3, which combined both expectations in a single design. Experiment 4 replicated and extended Experiment 3 to a broader set of conditions. Across both experiments, listeners based their inferences on both types of expectations. Overall, model predictions and data were highly correlated. Predictions from the model incorporating both types of information described listeners’ behavior better compared to alternative models based on only one type of information. This research shows that listeners flexibly integrate different forms of social expectations across a range of contexts, a process which can be described using Bayesian cognitive models. Our approach highlights how experiments and computational models can be used in conjunction to inform theories about language use and comprehension."

keywords:
    "Pragmatics; Word learning; Common ground; Bayesian models"

output: cogsci2016::cogsci_paper
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, fig.pos = "tb", fig.path='figs/',
                      echo=F, warning=F, cache=F, message=F, sanitize = T)
```

```{r, libraries}
library(png)
library(grid)
library(ggplot2)
library(xtable)
library(tidyverse)
library(knitr)
library(ggthemes)
library(langcog)
library(rwebppl)
library(matrixStats)
library(coda)
library(ggpubr)
library(lme4)
library(broom)
library(exactRankTests)
```

```{r}
# data ex 1 and ex 2
d1 <- bind_rows(read_csv(file="../../stats/data/novel.data.csv"),
               read_csv(file="../../stats/data/pref.data.csv")) %>%
  mutate(trial_type = ifelse(trial == "train", "train", "test"),
         control = ifelse(change == "true", "true" , "false"))

d2 <- read_csv(file="../../stats/data/inf.data.csv") %>%
  mutate(trial_type = ifelse(trial == "train", "train", "test"))

dex12 <- bind_rows(d1,d2) %>%
  filter(trial_type == "test")%>%
  mutate(control= relevel(factor(control), ref = "true"),
         choiceAgent = ifelse( control == "false", agent, altAgent)) # re-code agent for preference and novelty

# data ex 3
d3 <- bind_rows(read_csv(file="../../stats/data/ex3.novel.data.csv"),
               read_csv(file="../../stats/data/ex3.pref.data.csv")) %>%
  mutate(trial_type = ifelse(trial == "train1" | trial =="train2", "train", "test"))

# check if someone did both experiments
# yes, two people
x <- d3 %>%
  filter(trial_type == "train")%>%
  group_by(id)%>%
  summarise(n = length(correct_inf)) %>%
  filter(n > 2)

# exclude those who did both experiments
dex3 <- d3 %>%
  filter(!id %in% x$id) %>%
  mutate(Speaker = ifelse(change =="same", "Same speaker", "Different speaker"),
         Alignment = ifelse(alignment == "congruent", "Congruent", "Incongruent"),
         Prior = ifelse(experiment == "pref_inf", "Preference", "Novelty"),
         choiceAgent = ifelse (change == "same", agent, altAgent))%>%
  filter(trial_type == "test")

d4 <- bind_rows(read_csv(file="../../stats/data/ex3.2.novel.strong.data.csv"),
               read_csv(file="../../stats/data/ex3.2.novel.medium.data.csv"),
               read_csv(file="../../stats/data/ex3.2.novel.weak.data.csv"),
               read_csv(file="../../stats/data/ex3.2.pref.strong.data.csv"),
               read_csv(file="../../stats/data/ex3.2.pref.medium.data.csv")
               ) %>%
  mutate(trial_type = ifelse(trial == "train1" | trial =="train2", "train", "test"))

x <- d4 %>%
  filter(trial_type == "train") %>%
  group_by(id)%>%
  summarise(correct_inf = mean(correct_inf)) %>%
  filter(correct_inf == 0)

dex4 <- d4 %>%
  filter(!id %in% x$id)%>% 
  distinct(id, alltrial, .keep_all = TRUE) %>%
  filter(trial_type != "train") %>%
  mutate(Change = ifelse(change =="same", "Same speaker", "Different speaker"),
         Alignment = ifelse(alignment == "congruent", "Congruent", "Incongruent"),
         Experiment = ifelse(grepl("pref",experiment), "Preference", "Novelty"),
         Prior = ifelse(prior == "weak", "Weak", ifelse(prior == "medium", "Medium", "Strong"))
         )

# loading model comparisons and predictions

pred_ex3 <- readRDS(file = "../../stats/saves/model_predictions_ex3_1.rds")

pred_ex4 <- readRDS(file = "../../stats/saves/model_predictions_ex3_variable_priors.rds")

bf_ex3 <- readRDS(file = "../../stats/saves/ex3_bf.rds")

bf_ex4 <- readRDS(file = "../../stats/saves/ex4_bf.rds")

```

# Introduction

One of the most astonishing features of natural language is that it allows us to communicate precise meanings despite the fact that each utterance is inherently ambiguous. The conventional couplings between sounds (words) and objects constrain what a speaker may mean by an utterance. However, the intended meaning of the utterance is not reducible to the words that are contained in it. For example, the same utterance (“Look, there it is!”) could be intended to refer to very different things depending on whether someone is searching for their wallet or looking for the bus stop. It takes additional pragmatic inferences to recover the intended meaning [@levinson2000presumptive].

Pragmatic inferences rest on a set of expectations that interlocutors bring to the table when entering a communicative interaction. On the one hand, speakers and listeners have the general expectation that their partner communicates in an informative and relevant way [@sperber2001relevance]. Grice [-@grice1991studies] summarised this expectation via the cooperative principle: “Make your contribution such as is required, at the stage at which it occurs, by the accepted purpose or direction of the talk exchange in which you are engaged.” Importantly, the second half of the principle highlights a second type of expectation: interlocutors expect each other to produce and interpret utterances in light of the shared common ground between them [@clark1996using]. Common ground refers to bits of information that are assumed to be shared, either because they result from joint experience or were otherwise grounded [@bohn2018common]. The nature of the corresponding expectations vary with the individuals involved.

Both general and common ground expectations can support children’s word learning [@clark2009first; @tomasello2009constructing]. On the one hand, children have been found to learn novel words by assuming that the speaker is informative [@frank2014inferring]. That is, in the absence of any prior interaction with the speaker, children interpreted a novel word as referring to the most informative referent. On the other hand, children learn novel words by relying on common ground expectations [@akhtar1996role; @bohn2018common]. For example, when a speaker expressed preference for a particular object, children expect a novel word from the same speaker to refer to the previously preferred object [@saylor2009preschoolers].

But how are general and common ground expectations integrated with one another? Are pragmatic inferences strengthened additively when both support a particular interpretation? How are they weighed when they are in conflict? The rational speech act (RSA) framework [@frank2012predicting] offers a formal framework for this information integration problem. RSA models are characterized by their recursive structure in which speakers and listeners reason about each other’s interpretation of utterances in context. RSA models have been found to make accurate quantitative predictions about various forms of pragmatic language use and word learning [@goodman2016pragmatic]. However, a comprehensive treatment of how general and common ground expectations are integrated, is still missing.

@@

# Method

## General Design

Each experiment was implemented on a website (using JavaScript and HTML) to which participants were directed. Figure \ref{fig:design} provides a schematic overview of the setup and experimental procedures. The instructions informed participants that they will see a number of animal characters asking for novel toys. Participants’ task was to identify the toy a given animal is requesting. The basic layout involved two tables with toys on them, located left and right of a little hill, on which the animal was standing. For each animal, we recorded a set of utterances (one speaker per animal) that were used throughout the experiments to provide information and make requests. At test, toys were requested using the following utterance: “Oh cool, there is a [non-word] on the table, how neat, can you give me the [non-word]?”. Participants responded by clicking on one of the toys. Each experiment started with two training trials in which animals requested familiar objects (car and ball).

For all experiments, we pre-registered the sample size, experimental design and the statistical analysis. For Experiment 3 and 4, we also registered the model structure and predictions (see [https://osf.io/u7kxe](https://osf.io/u7kxe))

```{r design, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=5, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Schematic experimental procedure. In all conditions, at test (bottom), the speaker ambiguously requested an object using a non-word (e.g. “dax”). Participants clicked on the object they thought the speaker referred to. Informativeness (Experiment 1, left) translated to making one object less frequent in context. Common ground (Experiment 2, middle) was manipulated by making one object prefered by, or new to the speaker. Experiment 3 (right) combined manipulations. When expressing e.g. preference for an object on a table with two objects (panel 3 from top), the respective object was temporally enlarged. Condition shown here: preference - same speaker - incongruent."}
img <- png::readPNG("figs/method_figure.png")
grid::grid.raster(img)
```

# Experiment 1

## Participants, Design and Procedure

@@

## Results and Discussion

```{r ex1-results}
ex1t <- dex12 %>%
  filter(condition == "informativeness")%>%
  group_by(control, id) %>%
  summarise(correct = mean(correct)) %>%
  summarise(correct = list(correct)) %>%
  group_by(control) %>%
  mutate(df= t.test(unlist(correct), mu = 0.5)$parameter,
         t_value = t.test(unlist(correct), mu = 0.5)$statistic,
         p_value = t.test(unlist(correct), mu = 0.5)$p.value) %>%
  select(control,df,t_value,p_value)%>%
  mutate(p_value = ifelse(p_value<.001,"< .001",as.character(paste("=",substr(round(p_value,3),2,5),sep=" "))),
         t_value = round(t_value,2))

# model
lm_ex1 <- glmer(correct ~ control +
              (control |id) + (control |agent), 
              data = dex12%>%filter(condition == "informativeness"), 
              family = binomial,control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

ex1_r <- tidy(lm_ex1, effects = "fixed")%>%
  mutate(p.value = ifelse(p.value<.001,"< .001",as.character(paste("=",substr(round(p.value,3),2,5),sep=" "))),
         std.error = round(std.error,2),
         estimate = round(estimate,2))
```

Participants selected the less frequent object above chance in the test condition (t = `r ex1t%>%filter(control == "false")%>%pull(t_value)`, *p* `r ex1t%>%filter(control == "false")%>%pull(p_value)`, see Figure \ref{fig:plotexp12}) and did so more often compared to the control condition (genreralized linear mixed model (GLMM\footnote{All models included random effects for subject and speaker (animal).}):*$\beta$* = `r ex1_r%>%filter(term == "controlfalse")%>%pull(estimate)`, se = `r ex1_r%>%filter(term == "controlfalse")%>%pull(std.error)` *p*  `r ex1_r%>%filter(term == "controlfalse")%>%pull(p.value)`). This result replicates earlier work [@frank2014inferring].

```{r plotexp12, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3.4, fig.height=2, fig.cap = "Results from Experiment 1 and 2.For preference and novelty, control refers to a different speaker (see Fig. 1). Transparent dots show data from individual particpants, solid dots represent condition means, error bars are 95 \\% CIs. Dashed line indicates performance expected by chance."}

p1 <- dex12 %>%
  filter(trial_type == "test") %>%
  mutate(control = ifelse(control == "true", "Control", "Test"),
         condition = factor(condition), 
         experiment = ifelse(condition == "informativeness", "1","2"))%>%
  group_by(experiment,condition ,control, id) %>%
  summarise(correct = mean(correct))

p2 <- p1 %>%
  multi_boot_standard(col = "correct")

 ggplot() +
  geom_jitter(data = p1, aes(x = control, y = correct, col = control), alpha = 0.2, width = .3,height = .025)+
  geom_pointrange(data = p2, aes(x = control, y = mean, col = control,ymin = ci_lower, ymax = ci_upper),size = .7)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Expected Choice")+
  facet_wrap( ~ condition, labeller = as_labeller(c(`preference`="Exp. 2 - Pref.", `novelty`="Exp. 2 - Nov.", `informativeness`="Exp. 1 - Inf.")))+
  theme_few(base_size = 10) + 
  ylim(-0.05,1.05)+
  guides(alpha = F,size =F, col = F)+
  scale_color_manual(name = "Condition",
                     breaks=c("Test","Control"),
                     values= c("#839496","black"))

```

# Experiment 2

We manipulated common ground expectations based on two ways which have succesfully been used in developmental studies [e.g. @akhtar1996role; @saylor2009preschoolers]. Speakers either expressed preference for one object or one object was new to the speaker.

## Participants, Design and Procedure

There were 80 participants, 40 per manipulation. In the preference manipulation, each table had a different object. In the beginning, the animal appeared on the hill and introduced themselves. Next, they turned to one of the tables and expressed either that they liked (“Oh wow, I really like that one”) or disliked (“Oh bleh, I really don’t like that one”) the object before turning to the other side and expressing the respective other attitude. Then, the animal disappeared. After a short period of time, either the same or a different animal appeared and requested an object while facing straight (see Figure 1 in blue). If participants took into account the information they gained about the speaker, they should select the previously preferred object if the returning animal was the same. If a different animal returned, they should choose randomly between objects.

In the novelty manipulation, one table was initially empty while there was an object on the other one (see Figure \ref{fig:design}). The animal turned to one of the sides and commented either on the presence (“Aha, look at that”) or the absence of an object (“Hm, nothing there”). Next, the animal quickly disappeared. The same animal re-appeared and the sequence above was repeated. When the animal disappeared for the second time, a second object appeared on the empty table while the animal was away. Like in preference, we now manipulated if the same animal or a different one would return. In case of the same animal returning, listeners could infer the referent of the subsequent request by considering that one object was new to the speaker and therefore more likely to be of interest to them. No such inference was licensed when a different animal returned because both objects were equally new to them.

## Results and Discussion

```{r ex2-results-pref}
ex2tpref <- dex12 %>%
  filter(condition == "preference")%>%
  group_by(control, id) %>%
  summarise(correct = mean(correct)) %>%
  summarise(correct = list(correct)) %>%
  group_by(control) %>%
  mutate(df= t.test(unlist(correct), mu = 0.5)$parameter,
         t_value = t.test(unlist(correct), mu = 0.5)$statistic,
         p_value = t.test(unlist(correct), mu = 0.5)$p.value) %>%
  select(control,df,t_value,p_value)%>%
  mutate(p_value = ifelse(p_value<.001,"< .001",as.character(paste("=",substr(round(p_value,3),2,5),sep=" "))),
         t_value = round(t_value,2))

# model (maximally converging)
lm_ex2pref <- glmer(correct ~ control +
              (1|id) + (control |choiceAgent), 
              data = dex12%>%filter(condition == "preference"), 
              family = binomial,control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

ex2pref_r <- tidy(lm_ex2pref, effects = "fixed")%>%
  mutate(p.value = ifelse(p.value<.001,"< .001",as.character(paste("=",substr(round(p.value,3),2,5),sep=" "))),
         std.error = round(std.error,2),
         estimate = round(estimate,2))
```

```{r ex2-results-nov}
ex2tnov <- dex12 %>%
  filter(condition == "novelty")%>%
  group_by(control, id) %>%
  summarise(correct = mean(correct)) %>%
  summarise(correct = list(correct)) %>%
  group_by(control) %>%
  mutate(df= t.test(unlist(correct), mu = 0.5)$parameter,
         t_value = t.test(unlist(correct), mu = 0.5)$statistic,
         p_value = t.test(unlist(correct), mu = 0.5)$p.value) %>%
  select(control,df,t_value,p_value)%>%
  mutate(p_value = ifelse(p_value<.001,"< .001",as.character(paste("=",substr(round(p_value,3),2,5),sep=" "))),
         t_value = round(t_value,2))

# model
lm_ex2nov <- glmer(correct ~ control +
              (1|id) + (control |choiceAgent), 
              data = dex12%>%filter(condition == "novelty"), 
              family = binomial,control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

ex2nov_r <- tidy(lm_ex2nov, effects = "fixed")%>%
  mutate(p.value = ifelse(p.value<.001,"< .001",as.character(paste("=",substr(round(p.value,3),2,5),sep=" "))),
         std.error = round(std.error,2),
         estimate = round(estimate,2))
```

Participants selected the preferred object above chance when the same animal returned (t = `r ex2tpref%>%filter(control == "false")%>%pull(t_value)`, *p* `r ex2tpref%>%filter(control == "false")%>%pull(p_value)`, see Figure \ref{fig:plotexp12}) and did so more often compared to trials in which a different animal returned (GLMM: $\beta$ = `r ex2pref_r%>%filter(term == "controlfalse")%>%pull(estimate)`, se = `r ex2pref_r%>%filter(term == "controlfalse")%>%pull(std.error)`, *p* `r ex2pref_r%>%filter(term == "controlfalse")%>%pull(p.value)`). Thus, listeners inferred the referent of the utterance by considering previous interactions with the speaker. Interestingly, participants to some extend transferred preference from one animal to the other and also selected the preferred object above chance when a different animal returned (t = `r ex2tpref%>%filter(control == "true")%>%pull(t_value)`, *p* `r ex2tpref%>%filter(control == "true")%>%pull(p_value)`). In sum, this study shows that adults make comparable infernces to children [@saylor2009preschoolers].

The novel object was selected above chance when the same animal returned (t = `r ex2tnov%>%filter(control == "false")%>%pull(t_value)`, *p* `r ex2tnov%>%filter(control == "false")%>%pull(p_value)`) but not when a different one appeared (t = `r ex2tnov%>%filter(control == "true")%>%pull(t_value)`, *p* `r ex2tnov%>%filter(control == "true")%>%pull(p_value)`, see Figure \ref{fig:plotexp12}). Furthermore, the two conditions differed in the expected direction ($\beta$ = `r ex2nov_r%>%filter(term == "controlfalse")%>%pull(estimate)`, se = `r ex2nov_r%>%filter(term == "controlfalse")%>%pull(std.error)`, *p* `r ex2nov_r%>%filter(term == "controlfalse")%>%pull(p.value)`). Thus, like children [@akhtar1996role], participants used their prior information about the speaker to resolve ambiguity in their utterance.

# Experiment 3

```{r plotexp3, fig.env="figure*", fig.pos = "h", fig.width=7, fig.height=2.5, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Results from Experiment 3. Dashed line indicates performance expected by chance. Plotting conventions are the same as in Figure 2. All conditions in which CIs do not overlap with chance line are also statistically different from chance based on two-tailed Wilcoxon tests."}

p3 <- dex3 %>%
  group_by(Prior,Speaker ,Alignment, id) %>%
  summarise(correct = mean(correct_inf))

p4 <- p3 %>%
  multi_boot_standard(col = "correct")

ggplot() +
  #geom_rect(data = p4, aes(fill = NA), color = "#859900", size = 2.5, alpha = 0.4 ,xmin = -Inf,xmax = Inf, ymin = -Inf,ymax = Inf)+
  geom_jitter(data = p3, aes(x = Alignment, y = correct), alpha = 0.2, width = .3,height = .025)+
  geom_pointrange(data = p4, aes(x = Alignment, y = mean,ymin = ci_lower, ymax = ci_upper),size = .7)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Choosing More Informative")+
  facet_grid ( ~ Prior + Speaker) +
  theme_few(base_size = 10) + 
  ylim(-0.05,1.05)+
  guides(alpha = F, fill = F, col = F)+ 
  scale_fill_manual(name="Condition",
                     breaks=c("1","2","3"),
                     values= c("#859900","#859900","#859900"))

```

@@

## Participants, Design and Procedure

@@

## Model Predictions

To derive predictions, we used a probabilistic RSA model that simulates a pragmatic listener reasoning about a cooperative speaker who is trying to refer to an object [@frank2012predicting]. The speaker chooses how to refer to the object by reasoning about a naive listener who does not know the labels for the object [@frank2014inferring]. Mathematically, the probability that the listener chooses a referent given an utterance is defined as follows:  
$$P_L(r_s|u)\propto P_S(u|r_s)P_S(r_s)$$
Here, $P_S(u|r_s)$ is the likelihood that the speaker will use an utterance to refer to a specific referent. It is defined as the surprisal of an utterance given the referent for a naive listener, who interprets utterances in a literal way:

$$P_S(u|r_s)\propto exp(\alpha U_S(u;s))$$

The numerical strength of the expression above depends on $\alpha$, which can be interpreted as an indicator of how rational the speaker is in choosing utterances.

The term $P_S(r_s)$ denotes the prior probability that a speaker will refer to a given referent. This probability represents the listeners expectations about the speaker depending on the manipulation (preference or novelty) and the identity of the speaker (same or different speaker).

>>>>

% Template for Cogsci submission with R Markdown

% Stuff changed from original Markdown PLOS Template
\documentclass[10pt, letterpaper]{article}

\usepackage{cogsci}
\usepackage{pslatex}
\usepackage{float}
\usepackage{caption}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}

% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% hyperref package, useful for hyperlinks
\usepackage{hyperref}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% Sweave(-like)
\usepackage{fancyvrb}
\DefineVerbatimEnvironment{Sinput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{Soutput}{Verbatim}{}
\DefineVerbatimEnvironment{Scode}{Verbatim}{fontshape=sl}
\newenvironment{Schunk}{}{}
\DefineVerbatimEnvironment{Code}{Verbatim}{}
\DefineVerbatimEnvironment{CodeInput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{CodeOutput}{Verbatim}{}
\newenvironment{CodeChunk}{}{}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{color}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace}
%\doublespacing

% % Text layout
% \topmargin 0.0cm
% \oddsidemargin 0.5cm
% \evensidemargin 0.5cm
% \textwidth 16cm
% \textheight 21cm

\title{Integrating Common Ground and Informativeness in Pragmatic Word Learning}

\author{{\large \bf Manuel Bohn} \\ \texttt{bohn@stanford.edu} \\ Psychology, Stanford University \\ LFE, Leipzig University 
 \And {\large \bf Michael Henry Tessler} \\ \texttt{tessler@mit.edu} \\ Department of Brain and Cognitive Sciences \\ Massachusetts Institute of Technology
 \And {\large \bf Michael C. Frank} \\ \texttt{mcfrank@stanford.edu} \\ Department of Psychology \\ Stanford University}

\begin{document}

\maketitle

\begin{abstract}
Pragmatic inferences are an integral part of language comprehension and
learning. To recover the intended meaning of an utterance, listeners
need to balance and integrate different sources in information. In a
series of experiments, we studied how listeners integrate general
expectations about speakers with expectations that are specific to the
interactional history with a particular speaker. We used a Bayesian
cognitive model to formalize the integration process. In Experiment 1
and 2 we replicated findings showing that listeners make inferences
based on general and speaker specific expectations. Using the empirical
measurements from these experiments, we generated model predictions
about how they should be integrated. We tested these predictions in
Experiment 3, which combined both expectations in a single design.
Experiment 4 replicated and extended Experiment 3 to a broader set of
conditions. Across both experiments, listeners based their inferences on
both types of expectations. Overall, model predictions and data were
highly correlated. Predictions from the model incorporating both types
of information described listeners' behavior better compared to
alternative models based on only one type of information. This research
shows that listeners flexibly integrate different forms of social
expectations across a range of contexts, a process which can be
described using Bayesian cognitive models. Our approach highlights how
experiments and computational models can be used in conjunction to
inform theories about language use and comprehension.

\textbf{Keywords:}
Pragmatics; Word learning; Common ground; Bayesian models
\end{abstract}

\section{Introduction}\label{introduction}

One of the most astonishing features of natural language is that it
allows us to communicate precise meanings despite the fact that each
utterance is inherently ambiguous. The conventional couplings between
sounds (words) and objects constrain what a speaker may mean by an
utterance. However, the intended meaning of the utterance is not
reducible to the words that are contained in it. For example, the same
utterance (``Look, there it is!'') could be intended to refer to very
different things depending on whether someone is searching for their
wallet or looking for the bus stop. It takes additional pragmatic
inferences to recover the intended meaning (Levinson, 2000).

Pragmatic inferences rest on a set of expectations that interlocutors
bring to the table when entering a communicative interaction. On the one
hand, speakers and listeners have the general expectation that their
partner communicates in an informative and relevant way (Sperber \&
Wilson, 2001). Grice (1991) summarised this expectation via the
cooperative principle: ``Make your contribution such as is required, at
the stage at which it occurs, by the accepted purpose or direction of
the talk exchange in which you are engaged.'' Importantly, the second
half of the principle highlights a second type of expectation:
interlocutors expect each other to produce and interpret utterances in
light of the shared common ground between them (H. H. Clark, 1996).
Common ground refers to bits of information that are assumed to be
shared, either because they result from joint experience or were
otherwise grounded (Bohn \& Koymen, 2018). The nature of the
corresponding expectations vary with the individuals involved.

Both general and common ground expectations can support children's word
learning (E. V. Clark, 2009; Tomasello, 2009). On the one hand, children
have been found to learn novel words by assuming that the speaker is
informative (Frank \& Goodman, 2014). That is, in the absence of any
prior interaction with the speaker, children interpreted a novel word as
referring to the most informative referent. On the other hand, children
learn novel words by relying on common ground expectations (Akhtar,
Carpenter, \& Tomasello, 1996; Bohn \& Koymen, 2018). For example, when
a speaker expressed preference for a particular object, children expect
a novel word from the same speaker to refer to the previously preferred
object (Saylor, Sabbagh, Fortuna, \& Troseth, 2009).

But how are general and common ground expectations integrated with one
another? Are pragmatic inferences strengthened additively when both
support a particular interpretation? How are they weighed when they are
in conflict? The rational speech act (RSA) framework (Frank \& Goodman,
2012) offers a formal framework for this information integration
problem. RSA models are characterized by their recursive structure in
which speakers and listeners reason about each other's interpretation of
utterances in context. RSA models have been found to make accurate
quantitative predictions about various forms of pragmatic language use
and word learning (Goodman \& Frank, 2016). However, a comprehensive
treatment of how general and common ground expectations are integrated,
is still missing.

@@

\section{Method}\label{method}

\subsection{General Design}\label{general-design}

Each experiment was implemented on a website (using JavaScript and HTML)
to which participants were directed. Figure \ref{fig:design} provides a
schematic overview of the setup and experimental procedures. The
instructions informed participants that they will see a number of animal
characters asking for novel toys. Participants' task was to identify the
toy a given animal is requesting. The basic layout involved two tables
with toys on them, located left and right of a little hill, on which the
animal was standing. For each animal, we recorded a set of utterances
(one speaker per animal) that were used throughout the experiments to
provide information and make requests. At test, toys were requested
using the following utterance: ``Oh cool, there is a {[}non-word{]} on
the table, how neat, can you give me the {[}non-word{]}?''. Participants
responded by clicking on one of the toys. Each experiment started with
two training trials in which animals requested familiar objects (car and
ball).

For all experiments, we pre-registered the sample size, experimental
design and the statistical analysis. For Experiment 3 and 4, we also
registered the model structure and predictions (see
\url{https://osf.io/u7kxe})

\begin{CodeChunk}
\begin{figure*}[h]

{\centering \includegraphics{figs/design-1}

}

\caption[Schematic experimental procedure]{Schematic experimental procedure. In all conditions, at test (bottom), the speaker ambiguously requested an object using a non-word (e.g. “dax”). Participants clicked on the object they thought the speaker referred to. Informativeness (Experiment 1, left) translated to making one object less frequent in context. Common ground (Experiment 2, middle) was manipulated by making one object prefered by, or new to the speaker. Experiment 3 (right) combined manipulations. When expressing e.g. preference for an object on a table with two objects (panel 3 from top), the respective object was temporally enlarged. Condition shown here: preference - same speaker - incongruent.}\label{fig:design}
\end{figure*}
\end{CodeChunk}

\section{Experiment 1}\label{experiment-1}

\subsection{Participants, Design and
Procedure}\label{participants-design-and-procedure}

@@

\subsection{Results and Discussion}\label{results-and-discussion}

Participants selected the less frequent object above chance in the test
condition (t = 5.51, \emph{p} \textless{} .001, see Figure
\ref{fig:plotexp12}) and did so more often compared to the control
condition (genreralized linear mixed model
(GLMM\footnote{All models included random effects for subject and speaker (animal).}):\emph{\(\beta\)}
= 1.49, se = 0.5 \emph{p} = .003). This result replicates earlier work
(Frank \& Goodman, 2014).

\begin{CodeChunk}
\begin{figure}[H]

{\centering \includegraphics{figs/plotexp12-1}

}

\caption[Results from Experiment 1 and 2.For preference and novelty, control refers to a different speaker (see Fig]{Results from Experiment 1 and 2.For preference and novelty, control refers to a different speaker (see Fig. 1). Transparent dots show data from individual particpants, solid dots represent condition means, error bars are 95 \% CIs. Dashed line indicates performance expected by chance.}\label{fig:plotexp12}
\end{figure}
\end{CodeChunk}

\section{Experiment 2}\label{experiment-2}

We manipulated common ground expectations based on two ways which have
succesfully been used in developmental studies (e.g. Akhtar et al.,
1996; Saylor et al., 2009). Speakers either expressed preference for one
object or one object was new to the speaker.

\subsection{Participants, Design and
Procedure}\label{participants-design-and-procedure-1}

There were 80 participants, 40 per manipulation. In the preference
manipulation, each table had a different object. In the beginning, the
animal appeared on the hill and introduced themselves. Next, they turned
to one of the tables and expressed either that they liked (``Oh wow, I
really like that one'') or disliked (``Oh bleh, I really don't like that
one'') the object before turning to the other side and expressing the
respective other attitude. Then, the animal disappeared. After a short
period of time, either the same or a different animal appeared and
requested an object while facing straight (see Figure 1 in blue). If
participants took into account the information they gained about the
speaker, they should select the previously preferred object if the
returning animal was the same. If a different animal returned, they
should choose randomly between objects.

In the novelty manipulation, one table was initially empty while there
was an object on the other one (see Figure \ref{fig:design}). The animal
turned to one of the sides and commented either on the presence (``Aha,
look at that'') or the absence of an object (``Hm, nothing there'').
Next, the animal quickly disappeared. The same animal re-appeared and
the sequence above was repeated. When the animal disappeared for the
second time, a second object appeared on the empty table while the
animal was away. Like in preference, we now manipulated if the same
animal or a different one would return. In case of the same animal
returning, listeners could infer the referent of the subsequent request
by considering that one object was new to the speaker and therefore more
likely to be of interest to them. No such inference was licensed when a
different animal returned because both objects were equally new to them.

\subsection{Results and Discussion}\label{results-and-discussion-1}

Participants selected the preferred object above chance when the same
animal returned (t = 29.14, \emph{p} \textless{} .001, see Figure
\ref{fig:plotexp12}) and did so more often compared to trials in which a
different animal returned (GLMM: \(\beta\) = 2.97, se = 0.7, \emph{p}
\textless{} .001). Thus, listeners inferred the referent of the
utterance by considering previous interactions with the speaker.
Interestingly, participants to some extend transferred preference from
one animal to the other and also selected the preferred object above
chance when a different animal returned (t = 2.7, \emph{p} = .01). In
sum, this study shows that adults make comparable infernces to children
(Saylor et al., 2009).

The novel object was selected above chance when the same animal returned
(t = 6.77, \emph{p} \textless{} .001) but not when a different one
appeared (t = 1.49, \emph{p} = .144, see Figure \ref{fig:plotexp12}).
Furthermore, the two conditions differed in the expected direction
(\(\beta\) = 1.7, se = 0.39, \emph{p} \textless{} .001). Thus, like
children (Akhtar et al., 1996), participants used their prior
information about the speaker to resolve ambiguity in their utterance.

\section{Experiment 3}\label{experiment-3}

\begin{CodeChunk}
\begin{figure*}[h]

{\centering \includegraphics{figs/plotexp3-1}

}

\caption[Results from Experiment 3]{Results from Experiment 3. Dashed line indicates performance expected by chance. Plotting conventions are the same as in Figure 2. All conditions in which CIs do not overlap with chance line are also statistically different from chance based on two-tailed Wilcoxon tests.}\label{fig:plotexp3}
\end{figure*}
\end{CodeChunk}

@@

\subsection{Participants, Design and
Procedure}\label{participants-design-and-procedure-2}

@@

\subsection{Model Predictions}\label{model-predictions}

To derive predictions, we used a probabilistic RSA model that simulates
a pragmatic listener reasoning about a cooperative speaker who is trying
to refer to an object (Frank \& Goodman, 2012). The speaker chooses how
to refer to the object by reasoning about a naive listener who does not
know the labels for the object (Frank \& Goodman, 2014). Mathematically,
the probability that the listener chooses a referent given an utterance
is defined as follows:\\
\[P_L(r_s|u)\propto P_S(u|r_s)P_S(r_s)\] Here, \(P_S(u|r_s)\) is the
likelihood that the speaker will use an utterance to refer to a specific
referent. It is defined as the surprisal of an utterance given the
referent for a naive listener, who interprets utterances in a literal
way:

\[P_S(u|r_s)\propto exp(\alpha U_S(u;s))\]

The numerical strength of the expression above depends on \(\alpha\),
which can be interpreted as an indicator of how rational the speaker is
in choosing utterances.

The term \(P_S(r_s)\) denotes the prior probability that a speaker will
refer to a given referent. This probability represents the listeners
expectations about the speaker depending on the manipulation (preference
or novelty) and the identity of the speaker (same or different speaker).

In our model, we used the results from Experiment 1 and 2 to specify
\(\alpha\) as well as \(P_S(r_s)\). The former was set to yield the
proportion measured in Experiment 1 in a model with uniform priors. The
prior probability for each object was the proportion with which this
object was chosen in Experiment
2\footnote{Proportions were measured when participants chose between two objects. However, in Experiment 3, three objects were involved. For each object we used the proportion measured in Experiment 2 as the prior probability.This approached spread out the absolute probability mass for each object but conserved the relative relation between objects.}.
Based on these parameter settings, we predicted the proportion with
which listeners will choose the less frequent object in each unique
combination of alignment and speaker identity separately for each prior
manipulation (eight conditions total, see Figure \ref{fig:plotexp3}). We
compared the fit of this pragmatic model to two alternative models using
Bayes Factors (BF). The first alternative model ignored the speaker
specific expectations (flat prior model) while the second ignored the
informativeness inference (prior only model). To capture the idea that
behavioral data is to some degree noisy, all models included a ``noise
parameter'', which reflects that, with a certain probability,
participants respond randomly instead of in line with the intended
manipulation on a given trial. Noise parameters are estimated based on
the data.
